Carrier phase and amplitude estimation for phase shift keying using pilots and data

### 2012-12-29 paper.tex
added stuff

Figures~\ref{fig:plotphaseBPSK},~\ref{fig:plotphaseQPSK}~and~\ref{fig:plotphase8PSK} show the mean square error (MSE) of the phase estimator when $M=2,4,8$ with $L=4096$ and for varying proportions of pilots symbols $\abs{P} = 0, \tfrac{L}{2}, \tfrac{L}{8}, \tfrac{L}{32}, L$.  When $\abs{P} \neq 0$ (i.e. coherent detection) the mean square error is computed as $\tfrac{1}{T}\sum_{i=1}^T\sfracpart{\hat{\theta}_i - \theta_0}_\pi^2$.  Otherwise, when $\abs{P}=0$ the mean square error is computed as $\tfrac{1}{T}\sum_{i=1}^T\sfracpart{\hat{\theta}_i - \theta_0}^2$ as in~\cite{McKilliam_leastsqPSKnoncoICASSP_2012}.  The dots, squares, circles and crosses are the results of Monte-Carlo simulations with the least square estimator.  The solid lines are the MSEs predicted by Theorem~\ref{thm:normality}.  The theorem accurately predicts the behaviour of the phase estimator when the SNR is sufficiently large.  As the SNR decreases the variance of the phase estimator approaches that of the uniform distribution on $[-\pi, \pi)$ when $\abs{P} \neq 0$ and the uniform distribution on $[-\tfrac{\pi}{M}, \tfrac{\pi}{M})$ when $\abs{P}=0$~\cite{McKilliam_leastsqPSKnoncoICASSP_2012}.  Theorem~\ref{thm:normality} does not model this behaviour.

Figures~\ref{fig:plotampBPSK},~\ref{fig:plotampQPSK}~and~\ref{fig:plotamp8PSK} show the variance of the amplitude estimator when $M=2,4,8$ and with $L=32, 256, 2048$ and when the number of pilots symbols is $\abs{P} = 0, \tfrac{L}{2}, L$.  The solid lines are the variance predicted by Theorem~\ref{thm:normality}.  The dots and crosses show the results of Monte-Carlo simulations.  Each point is computed as the unbiased error $\tfrac{1}{T}\sum_{i=1}^T\big(\hat{\rho}_i - \rho_0G(0)\big)^2$.  This requires $G(0)$ to be known.  In practice $G(0)$ may not be known at the receiver, so Figures~\ref{fig:plotampBPSK},~\ref{fig:plotampQPSK}~and~\ref{fig:plotamp8PSK} serve to validate the correctness of our asymptotic theory, rather than to suggest the practical performance of the amplitude estimator.  When SNR is large $G(0)$ is close to $1$ and the bias of the amplitude estimator is small.  However, $G(0)$ grows without bound as the variance of the noise increases, so the bias is significant when SNR is small.  As indicated in Figures~\ref{fig:plotampBPSK}~and~\ref{fig:plotampQPSK}~and~\ref{fig:plotamp8PSK} the variance of the least squares amplitude estimator is smaller than that of the unmodulated carrier.  However, due to the bias, the MSE of the least squares amplitude estimator is \emph{not} smaller than that of the unmodulated carrier.

@@

\begin{lemma}\label{lem:HLtoG} $\sqrt{L}\big( T_L(\hat{\lambda}_L) - G(0) \big) = o_P(1)$.
\end{lemma}
\begin{IEEEproof}
The argument is similar to that used in Lemma~\ref{lem:Qconv}.  First observe that
\begin{align*}
T_L(\lambda) &= \frac{\sabs{P}}{L}\expect R_1\cos(\lambda + \Phi_1) + \frac{\sabs{D}}{L} \expect R_1\cos\fracpart{\lambda + \Phi_1} \\
%&= \big( p + o(1) \big) \expect R_1\cos(\lambda + \Phi_1) \\
%& \hspace{2cm} + \big( d + o(1) \big) \expect R_1\cos\fracpart{\lambda + \Phi_1} \\
&=  \big( p + o(L^{-1/2}) \big) \expect R_1\cos(\lambda + \Phi_1) \\
&\hspace{2cm} + \big( d + o(L^{-1/2}) \big) \expect R_1\cos\fracpart{\lambda + \Phi_1},
\end{align*}
and because $\tfrac{\abs{P}}{L} = p + o(L^{-1/2})$ and $\tfrac{\abs{D}}{L} = d + o(L^{-1/2})$ (by assumption in Theorem~\ref{thm:normality}), we have
\begin{align*}
\sqrt{L}\big( T_L(\lambda) - G(0) \big) = p q_1(\lambda) + d q_2(\lambda) + o(1),
\end{align*}
where
\begin{align}
q_1(\lambda) &= \expect R_1\big( \cos(\lambda + \Phi_1) - \cos(\Phi_1) \big) \;\;\; \text{and} \nonumber \\
q_2(\lambda) &= \expect R_1\big( \cos\fracpart{\lambda + \Phi_1} - \cos\fracpart{\Phi_1} \big). \label{eq:q2def}
\end{align}
Lemma~\ref{lem:q1k1parts} shows that $q_1(\hat{\lambda}_L) = \hat{\lambda}_Lo_P(1)$ and Lemma~\ref{lem:q2k2parts} shows that $q_2(\hat{\lambda}_L) = \hat{\lambda}_L o_P(1)$ and so
\[
\sqrt{L}\big( T_L(\lambda) - G(0) \big) =\sqrt{L}\hat{\lambda}_L o_P(1) + o(1).
\]
The lemma follows since $\sqrt{L}\hat{\lambda}_L$ converges in distribution.
\end{IEEEproof}

>>>>

Figures~\ref{fig:plotphaseBPSK},~\ref{fig:plotphaseQPSK}~and~\ref{fig:plotphase8PSK} show the mean square error (MSE) of the phase estimator when $M=2,4,8$ with $L=4096$ and for varying proportions of pilots symbols $\abs{P} = 0, \tfrac{L}{2}, \tfrac{L}{8}, \tfrac{L}{32}, L$.  When $\abs{P} \neq 0$ (i.e. coherent detection) the mean square error is computed as $\tfrac{1}{T}\sum_{i=1}^T\sfracpart{\hat{\theta}_i - \theta_0}_\pi^2$.  Otherwise, when $\abs{P}=0$ the mean square error is computed as $\tfrac{1}{T}\sum_{i=1}^T\sfracpart{\hat{\theta}_i - \theta_0}^2$ as in~\cite{McKilliam_leastsqPSKnoncoICASSP_2012}.  The dots, squares, circles and crosses are the results of Monte-Carlo simulations with the least square estimator.  The solid lines are the MSEs predicted by Theorem~\ref{thm:normality} (and by Theorem~2 from~\cite{McKilliam_leastsqPSKnoncoICASSP_2012} when $\abs{P} = 0$).  The theorem accurately predicts the behaviour of the phase estimator when the SNR is sufficiently large.  As the SNR decreases the variance of the phase estimator approaches that of the uniform distribution on $[-\pi, \pi)$ when $\abs{P} \neq 0$ and the uniform distribution on $[-\tfrac{\pi}{M}, \tfrac{\pi}{M})$ when $\abs{P}=0$~\cite{McKilliam_leastsqPSKnoncoICASSP_2012}.  Theorem~\ref{thm:normality} does not model this behaviour.

Figures~\ref{fig:plotampBPSK},~\ref{fig:plotampQPSK}~and~\ref{fig:plotamp8PSK} show the variance of the amplitude estimator when $M=2,4,8$ and with $L=32, 256, 2048$ and when the number of pilots symbols is $\abs{P} = 0, \tfrac{L}{2}, L$.  The solid lines are the variance predicted by Theorem~\ref{thm:normality}.  The dots and crosses show the results of Monte-Carlo simulations.  Each point is computed as the unbiased error $\tfrac{1}{T}\sum_{i=1}^T\big(\hat{\rho}_i - \rho_0G(0)\big)^2$.  This requires $G(0)$ to be known.  In practice $G(0)$ may not be known at the receiver, so Figures~\ref{fig:plotampBPSK},~\ref{fig:plotampQPSK}~and~\ref{fig:plotamp8PSK} serve to validate the correctness of our asymptotic theory, rather than to suggest the practical performance of the amplitude estimator.  When SNR is large $G(0)$ is close to $1$ and the bias of the amplitude estimator is small.  However, $G(0)$ grows without bound as the variance of the noise increases, so the bias is significant when SNR is small.  As indicated in Figures~\ref{fig:plotampBPSK},~\ref{fig:plotampQPSK}~and~\ref{fig:plotamp8PSK} the variance of the least squares amplitude estimator is smaller than that of the unmodulated carrier.  However, due to the bias, the MSE of the least squares amplitude estimator is \emph{not} smaller than that of the unmodulated carrier.

@@

\begin{lemma}\label{lem:HLtoG} $\sqrt{L}\big( T_L(\hat{\lambda}_L) - G(0) \big) = o_P(1)$.
\end{lemma}
\begin{IEEEproof}
The argument is similar to that used in Lemma~\ref{lem:Qconv}.  First observe that
\begin{align*}
T_L(\lambda) &= \frac{\sabs{P}}{L}\expect R_1\cos(\lambda + \Phi_1) + \frac{\sabs{D}}{L} \expect R_1\cos\fracpart{\lambda + \Phi_1} \\
%&= \big( p + o(1) \big) \expect R_1\cos(\lambda + \Phi_1) \\
%& \hspace{2cm} + \big( d + o(1) \big) \expect R_1\cos\fracpart{\lambda + \Phi_1} \\
&=  \big( p + o(L^{-1/2}) \big) \expect R_1\cos(\lambda + \Phi_1) \\
&\hspace{2cm} + \big( d + o(L^{-1/2}) \big) \expect R_1\cos\fracpart{\lambda + \Phi_1},
\end{align*}
and because $\tfrac{\abs{P}}{L} = p + o(L^{-1/2})$ and $\tfrac{\abs{D}}{L} = d + o(L^{-1/2})$ (by assumption in Theorem~\ref{thm:normality}), we have
\begin{align*}
\sqrt{L}\big( T_L(\lambda) - G(0) \big) = \sqrt{L}p q_1(\lambda) + \sqrt{L} d q_2(\lambda) + o(1),
\end{align*}
where
\begin{align}
q_1(\lambda) &= \expect R_1\big( \cos(\lambda + \Phi_1) - \cos(\Phi_1) \big) \;\;\; \text{and} \nonumber \\
q_2(\lambda) &= \expect R_1\big( \cos\fracpart{\lambda + \Phi_1} - \cos\fracpart{\Phi_1} \big). \label{eq:q2def}
\end{align}
Lemma~\ref{lem:q1k1parts} shows that $q_1(\hat{\lambda}_L) = \hat{\lambda}_Lo_P(1)$ and Lemma~\ref{lem:q2k2parts} shows that $q_2(\hat{\lambda}_L) = \hat{\lambda}_L o_P(1)$ and so
\[
\sqrt{L}\big( T_L(\lambda) - G(0) \big) =\sqrt{L}\hat{\lambda}_L o_P(1) + o(1).
\]
The lemma follows since $\sqrt{L}\hat{\lambda}_L$ converges in distribution.
\end{IEEEproof}
